Demand Forecasting of Individual Probability Density Functions with Machine Learning

### 2021-07-11 LaTex/paper.tex
some fixes from SNORF proof reads

@@

This work brings together elements from three different fields: operations research, machine learning, and causal inference. The combination of these allows to build state-of-the-art demand forecasting models that are fully explainable and incorporate the causal structure of demand. The contribution of this paper is two-fold:

@@

Furthermore, we show how predictions of full individual probability distributions, no matter if created by an approach with or without assumption of a specific distribution class, can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form (in the case of a parametric approach) as well as specific quantiles are predicted correctly. This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.

@@

Demand forecasting is one of the core operational activities of any customer-facing business and accurate demand predictions are vital to the optimization of the business operations. This is of particular concern for example for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}. It can be approached in a variety of ways, see, e.g., \cite{beheshti2015survey} for a survey of methods used in retail sales forecasting with a particular focus on fashion retail.

@@

% ML
Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached by means of supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.  For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting. Machine learning approaches can include multiple covariates. This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast. There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series. For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model. For example, we could add exponentially smoothed moving averages with a range of damping factors as features into the model. An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}. For the case of sequence samples, deep learning methods such as recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used. These methods employ information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events. Both concepts, LSTM cells and self-attention, aim at learning  the elements of the individual time series from the data that are deemed important for the prediction of the future values by the algorithm. The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}. A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}. A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.

@@

In the simplest case, we only have one model with one set of model parameters to cover all predictions. In this case, the evaluation of the full PDF is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as an NBD with predicted parameters, that is used for all observations. Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.

@@

For a first qualitative assessment, we make use of the probability integral transform, see e.g., \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1. We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution. Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

@@

In case of discrete probability distributions, such as the NBD, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g., \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items accounts for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$. Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g., for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.

@@

\fig \ref{fig:invquant_example} illustrates five different collections of inverse quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations. The dashed horizontal lines indicates the fraction we expect, i.e., the specified quantile value, if the predictions are correct. For example, for the median, the line at $q = 0.5$ indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line. The observation of the number of samples, indicated with different markers, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations. In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.

@@

% time series with ML and ewma features
Using machine learning, we express the task of time series forecasting as a supervised learning approach, where the prediction target, i.e., the quantity we wish to forecast, is represented by the values of the time series of interest itself. In the case of demand forecasting, the future expected demand is the variable we wish to forecast, and we need to build feature variables, either by hand or using some automatic approach, to construct a supervised machine learning model. A common approach is to exploit the autocorrelation of the target time series variable to build dedicated feature time series from its summary statistics of the respective recent past. For example, we could use an exponentially weighted moving average (EWMA), calculated according to \eqn \eqref{eqn:ewma}, for each individual time series and choose several different values for the smoothing parameter $\alpha$, which is defined in the range $[0, 1]$.
\begin{equation} \label{eqn:ewma}
\begin{split}
&\mathrm{EWMA}(x_{t+\mathrm{horizon}}) = \\ &\frac{x_t + (1 - \alpha) x_{t-1} + (1 - \alpha)^2 x_{t-2} + ... + (1 - \alpha)^t x_0}{1 + (1 - \alpha) + (1 - \alpha)^2 + ... + (1 - \alpha)^t}
\end{split}
\end{equation}
We can then use these EWMA features in the supervised machine learning model that we use to predict the future value of the target variable, in our case, the future demand for a given product at a given location. When constructing these features, we need to keep in mind that we use a lag with respect to the target time series corresponding to the forecast horizon. For example, if we were to forecast three days ahead into the future, the lag of the feature time series with respect to the target time series has to be three as well. By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past, and use these as features for the machine learning model. In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the products we wish to forecast, or characteristics of the location at which the product is sold.

@@

\noindent
The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the inverse dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values. The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl. The variance prediction for this day is driven by two competing factors: the Super Bowl event itself drives the variance up, while the high mean prediction (as feature) drives it down.

@@

\fig \ref{fig:cdf_demand} shows the histogram of CDF observations according to the method described in sec. \ref{sec:cdf_histo} for all product-location-day combinations in the test period. As benchmark, we compare the outcome of our negative binomial model to a simpler Poisson assumption, which has only a single model parameter, the mean. Using the same mean predictions for both negative binomial and Poisson model, the negative binomial PDF predictions are much closer to the uniform distribution, which we expect for optimal PDF predictions, than the Poisson PDF predictions, showing the effectiveness of our variance estimation. While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction). The first case is mainly due to a slight zero-inflation of actual sales. The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$. The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.

@@

\fig \ref{fig:invquant_mean} shows the inverse quantile profile plot using the predicted mean as the $x$ axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates. The considered quantiles are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}. While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model. For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above). And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter. The deviations at very low predicted mean values reflect the complications of zero-inflation and slow-sellers seen in the first and last bins of the CDF histograms above.

@@

In the case of the original newsvendor problem, underage and overage costs are linear. In the special case that they are identical, the cost function is given by the absolute deviation, i.e., $C(p,t) = |p-t|$.
We can derive the optimal point estimator in the following way, using the shorthand notation $\partial_p X$ for $\partial/\partial p X$
\begin{align*}
0 &= \partial_p E[E(p,t)]  \\
&= \partial_p \int_{-\infty}^\infty f(t) |p-t| dt \\
&= \partial_p \left ( \int_{-\infty}^{x_p} f(t) |p-t| dt +  \int_{x_p}^\infty f(t) |p-t| dt \right ), \\
\end{align*}
where we have split the integration at some intermediate point $x_q$. This allows us to resolve the absolute value if we choose $x_q$ such that $t$ is always smaller than $p$ in the first integral and the other way around in the second. Hence:
\begin{equation}
0 = \partial_p \left ( \int_{-\infty}^\infty f(t) (p-t) dt -  \int_{-\infty}^\infty f(t) (p-t) dt \right ) \\
\end{equation}
We then execute the partial derivative with respect to $p$ and obtain:
\begin{equation}
0 = -  \int_{-\infty}^{x_p} f(t) dt + \int_{x_p}^\infty f(t)  dt,
\end{equation}
which we can write as:
\begin{equation}
 \int_{-\infty}^{x_p} f(t) dt = \int_{x_p}^\infty f(t)  dt
\end{equation}
Therefore, the integral from $-\infty$ to the point $x_p$ is the same as from $x_p$ to $\infty$, which means that the point $x_p$ splits the integral in the interval $(-\infty, \infty)$ into two halves. This is the definition of the median, and therefore a linear cost function with equal over- and underage costs is linked to the median of the predicted PDF as optimal point estimator. If the underage and overage costs differ, the slopes of the two linear segments are no longer the same. This means that the resulting optimal point estimator is no longer the median, but a different quantile.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{profile}
\caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}
\end{center}
\end{figure}

>>>>

{DOI: 10.1007/s43069-021-00079-8}

{Published in Springer Operations Research Forum.}

@@

This work brings together elements from three different fields: operations research, machine learning, and causal inference. The combination of these allows to build state-of-the-art demand forecasting models that are fully explainable and incorporate the causal structure of demand. The contribution of this paper is twofold:

@@

Furthermore, we show how predictions of full individual probability distributions, no matter if created by an approach with or without assumption of a specific distribution class, can be evaluated qualitatively and quantitatively, such that we can assess whether both overall functional form (in the case of a parametric approach) as well as specific quantiles are predicted correctly. This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.

@@

Demand forecasting is one of the core operational activities of any customer-facing business and accurate demand predictions are vital to the optimization of the business operations. This is of particular concern, for example, for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}. It can be approached in a variety of ways, see, e.g., \cite{beheshti2015survey} for a survey of methods used in retail sales forecasting with a particular focus on fashion retail.

@@

% ML
Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached by means of supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.  For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting. Machine learning approaches can include multiple covariates. This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast. There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series. For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model. For example, we could add exponentially smoothed moving averages with a range of damping factors as features into the model. An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}. For the case of sequence samples, deep learning methods such as recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used. These methods employ information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events. Both concepts, LSTM cells and self-attention, aim at learning  the elements of the individual time series from the data that are deemed important for the prediction of the future values by the algorithm. The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}. A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, and the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}. A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.

@@

In the simplest case, we only have one model with one set of model parameters to cover all predictions. In this case, the evaluation of the full PDF is straightforward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as an NBD with predicted parameters, that is used for all observations. Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.

@@

For a first qualitative assessment, we make use of the probability integral transform, see, e.g., \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1. We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution. Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.

@@

In case of discrete probability distributions, such as the NBD, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g., \cite[p. 54]{casella2002statistical}. In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF. For example, if the sales of zero items account for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$. Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g., for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.

@@

\fig \ref{fig:invquant_example} illustrates five different collections of inverse quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations. The dashed horizontal lines indicate the fraction we expect, i.e., the specified quantile value, if the predictions are correct. For example, for the median, the line at $q = 0.5$ indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line. The observation of the number of samples, indicated with different markers, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations. In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.

@@

% time series with ML and ewma features
Using machine learning, we express the task of time series forecasting as a supervised learning approach, where the prediction target, i.e., the quantity we wish to forecast, is represented by the values of the time series of interest itself. In the case of demand forecasting, the future expected demand is the variable we wish to forecast, and we need to build feature variables, either by hand or using some automatic approach, to construct a supervised machine learning model. A common approach is to exploit the autocorrelation of the target time series variable to build dedicated feature time series from its summary statistics of the respective recent past. For example, we could use an exponentially weighted moving average (EWMA), calculated according to \eqn \eqref{eqn:ewma}, for each individual time series and choose several different values for the smoothing parameter $\alpha$, which is defined in the range $[0, 1]$.
\begin{equation} \label{eqn:ewma}
\begin{split}
&\mathrm{EWMA}(x_{t+\mathrm{horizon}}) = \\ &\frac{x_t + (1 - \alpha) x_{t-1} + (1 - \alpha)^2 x_{t-2} + ... + (1 - \alpha)^t x_0}{1 + (1 - \alpha) + (1 - \alpha)^2 + ... + (1 - \alpha)^t}
\end{split}
\end{equation}
We can then use these EWMA features in the supervised machine learning model that we use to predict the future value of the target variable, in our case, the future demand for a given product at a given location. When constructing these features, we need to keep in mind that we use a lag with respect to the target time series corresponding to the forecast horizon. For example, if we were to forecast three days ahead into the future, the lag of the feature time series with respect to the target time series has to be three as well. By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past and use these as features for the machine learning model. In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the products we wish to forecast, or characteristics of the location at which the product is sold.

@@

\noindent
The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the inverse dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends toward lower values. The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl. The variance prediction for this day is driven by two competing factors: the Super Bowl event itself drives the variance up, while the high mean prediction (as feature) drives it down.

@@

\fig \ref{fig:cdf_demand} shows the histogram of CDF observations according to the method described in sec. \ref{sec:cdf_histo} for all product-location-day combinations in the test period. As benchmark, we compare the outcome of our negative binomial model to a simpler Poisson assumption, which has only a single model parameter, the mean. Using the same mean predictions for both negative binomial and Poisson model, the negative binomial PDF predictions are much closer to the uniform distribution, which we expect for optimal PDF predictions, than the Poisson PDF predictions, showing the effectiveness of our variance estimation. While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction). The first case is mainly due to a slight zero-inflation of actual sales. The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$. The resulting distribution for the negative binomial model indicates a slight tendency toward too broad PDF predictions, at least for some observations.

@@

\fig \ref{fig:invquant_mean} shows the inverse quantile profile plot using the predicted mean as the $x$ axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates. The considered quantiles are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}. While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model. For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing toward too broad PDF predictions (as also observed in the CDF histogram above). And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter. The deviations at very low predicted mean values reflect the complications of zero-inflation and slow-sellers seen in the first and last bins of the CDF histograms above.

@@

In the case of the original newsvendor problem, underage and overage costs are linear. In the special case that they are identical, the cost function is given by the absolute deviation, i.e., $C(p,t) = |p-t|$.
We can derive the optimal point estimator in the following way, using the shorthand notation $\partial_p X$ for $\partial/\partial p X$
\begin{align*}
0 &= \partial_p E[E(p,t)]  \\
&= \partial_p \int_{-\infty}^\infty f(t) |p-t| dt \\
&= \partial_p \left ( \int_{-\infty}^{x_p} f(t) |p-t| dt +  \int_{x_p}^\infty f(t) |p-t| dt \right ), \\
\end{align*}
where we have split the integration at some intermediate point $x_q$. This allows us to resolve the absolute value if we choose $x_q$ such that $t$ is always smaller than $p$ in the first integral and the other way around in the second. Hence:
\begin{equation}
0 = \partial_p \left ( \int_{-\infty}^\infty f(t) (p-t) dt -  \int_{-\infty}^\infty f(t) (p-t) dt \right ) \\
\end{equation}
We then execute the partial derivative with respect to $p$ and obtain:
\begin{equation}
0 = -  \int_{-\infty}^{x_p} f(t) dt + \int_{x_p}^\infty f(t)  dt,
\end{equation}
which we can write as:
\begin{equation}
 \int_{-\infty}^{x_p} f(t) dt = \int_{x_p}^\infty f(t)  dt
\end{equation}
Therefore, the integral from $-\infty$ to the point $x_p$ is the same as from $x_p$ to $\infty$, which means that the point $x_p$ splits the integral in the interval $(-\infty, \infty)$ into two halves. This is the definition of the median, and therefore, a linear cost function with equal over- and underage costs is linked to the median of the predicted PDF as optimal point estimator. If the underage and overage costs differ, the slopes of the two linear segments are no longer the same. This means that the resulting optimal point estimator is no longer the median, but a different quantile.

@@

\begin{figure}
\begin{center}
\includegraphics[width=4cm]{profile}
\caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}
\end{center}
\end{figure}
